---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# BoltzMM

The BoltzMM package allows for computation of probability mass functions of fully-visible Boltzmann machines via \code{pfvbm} and \code{allpfvbm}.
Random data can be generated using \code{rfvbm}. Maximum pseudolikelihood estimation of parameters via the MM algorithm can be conducted using \code{fitfvbm}.
Computation of partial derivatives and Hessians can be performed via \code{fvbmpartiald} and \code{fvbmHessian}.
Covariance estimation and normal standard errors can be computed using \code{fvbmcov} and \code{fvbmstderr}.

## Installation

You can install BoltzMM from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("andrewthomasjones/BoltzMM")
```

## Examples

Compute the probability of every length n=3 binary spin vector under bvec and Mmat:

```{r example, eval = FALSE}
bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
allpfvbm(bvec,Mmat)
```

Generate num=1000 random strings of n=3 binary spin variables under bvec and Mmat.
```{r example2, eval = FALSE}
num <- 1000
bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
data <- rfvbm(num,bvec,Mmat)
```

Fit a fully visible Boltzmann machine to data, starting from parameters bvec and Mmat.
 
```{r example3, eval = FALSE}
fitfvbm(data,bvec,Mmat)
```

For more examples see individual help files.
