---
output: github_document
---

<img src="http://www.r-pkg.org/badges/version-last-release/BoltzMM"></img></a>
[![Downloads from the RStudio CRAN mirror](http://cranlogs.r-pkg.org/badges/BoltzMM)](https://CRAN.R-project.org/package=BoltzMM)
[![Build Status](https://travis-ci.org/andrewthomasjones/BoltzMM.svg?branch=master)](https://travis-ci.org/andrewthomasjones/BoltzMM)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# BoltzMM

The BoltzMM package allows for computation of probability mass functions of fully-visible Boltzmann machines via `pfvbm` and `allpfvbm`.
Random data can be generated using `rfvbm`. Maximum pseudolikelihood estimation of parameters via the MM algorithm can be conducted using `fitfvbm`.
Computation of partial derivatives and Hessians can be performed via `fvbmpartiald` and `fvbmHessian`.
Covariance estimation and normal standard errors can be computed using `fvbmcov` and `fvbmstderr`.

## Installation

You can install BoltzMM from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("andrewthomasjones/BoltzMM")
```

## Examples

Compute the probability of every length n=3 binary spin vector under bvec and Mmat:

```{r example, eval = T}
library(BoltzMM)
set.seed(1)

bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
allpfvbm(bvec,Mmat)
```

Generate num=1000 random strings of n=3 binary spin variables under bvec and Mmat.
```{r example2, eval = T}
library(BoltzMM)
set.seed(1)

num <- 1000
bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
data <- rfvbm(num,bvec,Mmat)

head(data)
```

Fit a fully visible Boltzmann machine to data, starting from parameters bvec and Mmat.
 
```{r example3, eval = T}
library(BoltzMM)
set.seed(1)

bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
data <- rfvbm(num,bvec,Mmat)

fitfvbm(data,bvec,Mmat)
```

For more examples see individual help files.
