---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


<img src="http://www.r-pkg.org/badges/version-last-release/BoltzMM"></img></a>
[![Downloads from the RStudio CRAN mirror](http://cranlogs.r-pkg.org/badges/BoltzMM)](https://CRAN.R-project.org/package=BoltzMM)
[![Build Status](https://travis-ci.org/andrewthomasjones/BoltzMM.svg?branch=master)](https://travis-ci.org/andrewthomasjones/BoltzMM)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# BoltzMM

The BoltzMM package allows for computation of probability mass functions of fully-visible Boltzmann machines via `pfvbm` and `allpfvbm`.
Random data can be generated using `rfvbm`. Maximum pseudolikelihood estimation of parameters via the MM algorithm can be conducted using `fitfvbm`.
Computation of partial derivatives and Hessians can be performed via `fvbmpartiald` and `fvbmHessian`.
Covariance estimation and normal standard errors can be computed using `fvbmcov` and `fvbmstderr`.

## Installation

If `devtools` has already been installed, then the most current build of `BoltzMM` can be obtained via the command:

```{r GH-install, eval=F}
devtools::install_github('andrewthomasjones/BoltzMM',build_vignettes = T)
```

The latest stable build of `BoltzMM` can be obtain from CRAN via the command:

```{r CRAN-install, eval = F}
install.packages("BoltzMM", repos='http://cran.us.r-project.org')
```

An archival build of `BoltzMM` is available at [url]. Manual installation instructions can be found within the *R* installation and administration manual https://cran.r-project.org/doc/manuals/r-release/R-admin.html.

## Examples

Compute the probability of every length n=3 binary spin vector under bvec and Mmat:

```{r example, eval = T}
library(BoltzMM)
set.seed(1)

bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
allpfvbm(bvec,Mmat)
```

Generate num=1000 random strings of n=3 binary spin variables under bvec and Mmat.
```{r example2, eval = T}
library(BoltzMM)
set.seed(1)

num <- 1000
bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
data <- rfvbm(num,bvec,Mmat)

head(data)
```

Fit a fully visible Boltzmann machine to data, starting from parameters bvec and Mmat.
 
```{r example3, eval = T}
library(BoltzMM)
set.seed(1)

bvec <- c(0,0.5,0.25)
Mmat <- matrix(0.1,3,3) - diag(0.1,3,3)
data <- rfvbm(num,bvec,Mmat)

fitfvbm(data,bvec,Mmat)
```

Example with real data from https://hal.archives-ouvertes.fr/hal-01927188v1.
```{r exampleReal, eval = T}
# Load bnstruct library & package
library(bnstruct)
library(BoltzMM)

#load data
data(senate)

# Turn data into a matrix
senate_data <- as.matrix(senate)

#recode Yes as 1, and No as -1
senate_data[senate_data=="Yes"]<-1
senate_data[senate_data=="No"]<--1

# Conduct imputation for missing data
imp_data <- knn.impute(senate_data,k=1)

# No governement - using as reference level
data_nogov <- imp_data[,-1]


# Initialize parameters
bvec <- rep(0,8)
Mmat <- matrix(0,8,8)
nullmodel<-list(bvec=bvec,Mmat=Mmat)

# Fit a fully visible Boltzmann machine to data, starting from parameters bvec and Mmat.
model <- fitfvbm(data_nogov,bvec,Mmat)
# Compute the sandwich covariance matrix using the data and the model.
covarmat <- fvbmcov(data_nogov,model,fvbmHess)
# Compute the standard errors of the parameter elements according to a normal approximation.
st_errors <- fvbmstderr(data_nogov,covarmat)
# Compute z-scores and p-values under null
test_results<-fvbmtests(data_nogov,model,nullmodel)

test_results
```

For more examples see individual help files.

## Unit testing

Using the package `testthat`, we have conducted the following unit test for the GitHub build, on the date: `r format(Sys.time(), '%d %B, %Y')`. The testing files are contained in the [tests](https://github.com/andrewthomasjones/BoltzMM/tree/master/tests) folder of the respository.

```{r unittest,  eval = F}

## Load 'BoltzMM' library.
library(BoltzMM)
  
## Load 'testthat' library.
library(testthat)

## Test 'BoltzMM'
test_package("BoltzMM")
```

## Bug reporting and contributions

Thank you for your interest in `BoltzMM`. If you happen to find any bugs in the program, then please report them on the Issues page (https://github.com/andrewthomasjones/BoltzMM/issues). Support can also be sought on this page. Furthermore, if you would like to make a contribution to the software, then please forward a pull request to the owner of the repository.

